---
title: " Modern Data Mining, HW 1"
author:
- Group Member 1
- Group Member 2
- Group Member 3
date: 'Due: 11:59PM,  Jan. 29th, 2023'
output:
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "hide", fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output
# check if you have ISLR package, if not, install it
if(!require('pacman')) {install.packages('pacman')}
pacman::p_load(ISLR, readxl, tidyverse, magrittr, dplyr, ggplot2,readxl,reshape2, ggpmisc)
```




\pagebreak

# Overview

This is a fast-paced course that covers a lot of material. There will be a large amount of references. You may need to do your own research to fill in the gaps in between lectures and homework/projects. It is impossible to learn data science without getting your hands dirty. Please budget your time evenly. Last-minute work ethic will not work for this course. 

Homework in this course is different from your usual homework assignment as a typical student. Most of the time, they are built over real case studies.  While you will be applying methods covered in lectures, you will also find that extra teaching materials appear here.  The focus will be always on the goals of the study, the usefulness of the data gathered, and the limitations in any conclusions you may draw. Always try to challenge your data analysis in a critical way. Frequently, there are no unique solutions. 

Case studies in each homework can be listed as your data science projects (e.g. on your CV) where you see fit. 



## Objectives 

- Get familiar with `R-studio` and `RMarkdown`
- Hands-on R 
- Learn data science essentials 
    - gather data
    - clean data
    - summarize data 
    - display data
    - conclusion
- Packages
    - `dplyr`
    - `ggplot`

##  Instructions

- **Homework assignments can be done in a group consisting of up to three members**. Please find your group members as soon as possible and register your group on our Canvas site.

- **All work submitted should be completed in the R Markdown format.** You can find a cheat sheet for R Markdown [here](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) For those who have never used it before, we urge you to start this homework as soon as possible. 

- **Submit the following files, one submission for each group:**  (1) Rmd file, (2) a compiled  HTML or pdf version, and (3) all necessary data files if different from our source data. You may directly edit this .rmd file to add your answers. If you intend to work on the problems separately within your group, compile your answers into one Rmd file before submitting. We encourage that you at least attempt each problem by yourself before working with your teammates. Additionally, ensure that you can 'knit' or compile your Rmd file. It is also likely that you need to configure Rstudio to properly convert files to PDF. [**These instructions**](http://kbroman.org/knitr_knutshell/pages/latex.html#converting-knitrlatex-to-pdf) might be helpful.

- In general, be as concise as possible while giving a fully complete answer to each question. All necessary datasets are available in this homework folder on Canvas. Make sure to document your code with comments (written on separate lines in a code chunk using a hashtag `#` before the comment) so the teaching fellows can follow along. R Markdown is particularly useful because it follows a 'stream of consciousness' approach: as you write code in a code chunk, make sure to explain what you are doing outside of the chunk. 

- A few good or solicited submissions will be used as sample solutions. When those are released, make sure to compare your answers and understand the solutions.


## Review materials

- Study Basic R Tutorial
- Study Advanced R Tutorial (to include `dplyr` and `ggplot`)
- Study lecture 1: Data Acquisition and EDA


# Case study 1: Audience Size

How successful is the Wharton Talk Show [Business Radio Powered by the Wharton School](https://businessradio.wharton.upenn.edu/)  


**Background:** Have you ever listened to [SiriusXM](https://www.siriusxm.com/)? Do you know there is a **Talk Show** run by Wharton professors in Sirius Radio?  Wharton launched a talk show called [Business Radio Powered by the Wharton School](https://businessradio.wharton.upenn.edu/) through the Sirius Radio station in January of 2014. Within a short period of time the general reaction seemed to be overwhelmingly positive. To find out the audience size for the show, we designed a survey and collected a data set via MTURK in May of 2014. Our goal was to **estimate the audience size**. There were 51.6 million Sirius Radio listeners then. One approach is to estimate the proportion of the Wharton listeners to that of the Sirius listeners, $p$, so that we will come up with an audience size estimate of approximately 51.6 million times $p$. 

To do so, we launched a survey via Amazon Mechanical Turk ([MTurk](https://www.mturk.com/)) on May 24, 2014 at an offered price of \$0.10 for each answered survey.  We set it to be run for 6 days with a target maximum sample size of 2000 as our goal. Most of the observations came in within the first two days. The main questions of interest are "Have you ever listened to Sirius Radio" and "Have you ever listened to Sirius Business Radio by Wharton?". A few demographic features used as control variables were also collected; these include Gender, Age and Household Income.  

We requested that only people in United States answer the questions. Each person can only fill in the questionnaire once to avoid duplicates. Aside from these restrictions, we opened the survey to everyone in MTurk with a hope that the sample would be more randomly chosen. 

The raw data is stored as `Survey_results_final.csv` on Canvas.


## Data preparation

1. We need to clean and select only the variables of interest. 

Select only the variables Age, Gender, Education Level, Household Income in 2013, Sirius Listener?, Wharton Listener? and Time used to finish the survey.

Reading in the data 
```{r}
setwd("./data")
survey = read.csv("Survey_results_final.csv")
```
```{r}
names(survey)
```

```{r}
survey.sub = survey %>%
  select(Answer.Age, Answer.Gender, Answer.Education, Answer.HouseHoldIncome , Answer.Sirius.Radio, Answer.Wharton.Radio, WorkTimeInSeconds)
```



Change the variable names to be "age", "gender", "education", "income", "sirius", "wharton", "worktime".


```{r}
colnames(survey.sub) = c("age", "gender", "education", "income", "sirius", "wharton", "worktime")
survey.sub
```


2. Handle missing/wrongly filled values of the selected variables

As in real world data with user input, the data is incomplete, with missing values, and has incorrect responses. There is no general rule for dealing with these problems beyond “use common sense.” In whatever case, explain what the problems were and how you addressed them. Be sure to explain your rationale for your chosen methods of handling issues with the data. Do not use Excel for this, however tempting it might be.

Tip: Reflect on the reasons for which data could be wrong or missing. How would you address each case? For this homework, if you are trying to predict missing values with regression, you are definitely overthinking. Keep it simple.

```{r}
missing.survey = survey.sub %>%
  summarise(nas = sum(is.na(survey.sub)),
            blanks = sum(survey.sub == ""),
            select = sum(survey.sub =="select one"))
missing.survey
```


How many blanks in
```{r}
# change value of wrong age values

blanks =  survey.sub %>%
  filter(age == "" | gender == "" | education == "" | education == "select one" | income == ""| sirius == "" | wharton == "" | worktime == "")

head(blanks)
```


b
Before this point, I thought  there were just blanks and NAs in the data, but discovered there were unchanged entire fields in the form os "select one" for at least the eucatuon column. I went back to tabulate these entires with `summarise()`. The reason the table shows 41 total counts at the filer shows 34 rows is because some people have multiple variables missing. 


This shows there aren't any NAs in the data, but there are 22 "blanks" or missing values. There are also more unanswered questions which indicate the "select one" option that people have not changed. 

Income could be missing because people did not have an income in the indicated year. Education could be missing if people have little or did not know what to report. 

Altogehter there are 34 entries which match this condition and will be reomved be removed. 

```{r}
dim(survey.sub)
age <- survey.sub$age 
age[age == "Eighteen (18)"] <-  18
age <- as.numeric((age))

survey.sub$age <- age
survey.sub.pruned = survey.sub %>%
  filter(!is.na(age) & !gender == "" & !education == "" & !education == "select one" & !income == ""& !sirius == "" & !wharton == "" & !worktime == "")

dim(survey.sub.pruned)

```

Tip: Reflect on the reasons for which data could be wrong or missing. How would you address each case? For this homework, if you are trying to predict missing values with regression, you are definitely overthinking. Keep it simple.


* Within the `age` variable, 3 people did not respond, 2 people selected ages that did not make sense in the context of the survey (4 and 223), and one person wrote their age as a character "eighteen (18)". Since these were likely errors on the part of the participant, we decided to fully remove their surveys

```{r examine age, echo=FALSE}
# coerce to numeric 
age <- survey.sub$age 
age[age == "Eighteen (18)"] <-  18
age <- as.numeric((age))
summary(age)

# 4 people did not respond and 2 people selected an incorrect age, remove these responses

```
* Within the `gender` variable, 6 people did not respond. We decided to fully remove their surveys.

```{r examine gender}
# coerce to factor 
survey.sub$gender %>% as.factor(.) %>% summary(.)

# 6 people did not respond, will remove these
```

* Within the `education` variable, 19 selected 'select one'. We decided to keep their surveys but change the result to "other".
```{r examine education}
education <- survey.sub$education %>% as.factor(.)
summary(education)

# 19 people selected the invalid response "select one" will chage these responses to 'other'
summary(education)
```

* Within the `income` variable, 6 people did not respond. We decided to fully remove their surveys.
```{r examine income}
# coerce to factor 
survey.sub$income %>% as.factor(.) %>% summary(.)

# 6 people did not respond, will remove these
```
* Within the `sirius` variable, 5 people did not respond. We decided to fully remove their surveys.
```{r examine sirius}
# coerce to factor 
survey.sub$sirius %>% as.factor(.) %>% summary(.)
```

* Within the `wharton` variable, 4 people did not respond. We decided to fully remove their surveys.
```{r examine wharton}
# coerce to factor 
survey.sub$wharton %>% as.factor(.) %>% summary(.)

# remove everyone who did not answer this question
# remove everyone who said no to sirius and yes to wharton
```


* We did not remove any values from the `worktime` variable
```{r examine worktime}
survey.sub$worktime %>% summary(.)

# this section is fine and doesn't need change
```


3. Brief summary 

Write a brief report to summarize all the variables collected. Include both summary statistics (including sample size) and graphical displays such as histograms or bar charts where appropriate. Comment on what you have found from this sample. (For example - it's very interesting to think about why would one work for a job that pays only 10cents/each survey? Who are those survey workers? The answer may be interesting even if it may not directly relate to our goal.)

received 1740 surveys
- gender
  - 42% Female, 58% male
- age 
  - right skewed, younger participants 
  - youngest was 18 and the oldest was 76
  - most people were in the early to late twenties
- Income
  - 11.90% Less than 15,000
  - 20.86% 15,000 - 30,000
  - 24.37% $30,000 - 50,000
  - 21.4% 50,000 - 75,000
  - 18.79% 75,000 - 150,000
  - 2.46% Above $150,000
- education 
  - 1.092% Other
  - 0.575% Less than 12 years; no high school diploma
  - 10.747% High school graduate (or equivalent)
  - 42.299% Some college, no diploma; or Associate’s degree
  - 35.115% Bachelor’s degree or other 4-year degree
  - 10.172% Graduate or professional degree
- worktime
  - average worktime 22.5 seconds
  - longest time 108 seconds
  - shortest time 8 seconds


```{r cleaned data}

# refactor education responses
survey_data_subset$education <- education
print(paste("number of surveys:", nrow(survey_data_subset)))
survey_data_cleaned <- filter(survey_data_subset,
                              (!is.na(age) & !(age %in% c(4,223))) &
                                gender != "" &
                                income != "" &
                                wharton != "" &
                                sirius != "" &
                                !(sirius == "No" & wharton == "Yes"))
print(paste("number of surveys:", nrow(survey_data_cleaned)))

survey_data_cleaned$income <- as.factor(survey_data_cleaned$income)
summary(survey_data_cleaned$income)

survey_data_cleaned$education <- as.factor(survey_data_cleaned$education)
summary(survey_data_cleaned$education)

survey_data_cleaned$gender <- as.factor(survey_data_cleaned$gender)
summary(survey_data_cleaned$gender)
```

```{r}

```


```{r gender plots}

# plot looking at gender
survey_data_cleaned %>%
  summarise(pct = summary(gender)/length(gender),
            gender = levels(gender)) %>%
  ggplot(aes(x=gender, y=pct, fill=pct))+
  geom_bar(stat= "identity")  +
  geom_text(aes(label = scales::percent(pct)), vjust = -.5) +
  labs(title = "Participants by Gender", y = "Percent", x ='Gender') +
  theme_minimal()

```



```{r age plots}

summary(survey_data_cleaned$age)

# plot looking at age total
ggplot(survey_data_cleaned) +
  geom_histogram( aes(x=age, y = ..density..),bins=30, fill= "forestgreen", position="identity") 

ggplot(survey_data_cleaned) +
  geom_boxplot(aes(y=age, x= gender, fill=gender),
                  position="identity") +
  scale_fill_brewer(palette="Reds")

ggplot(survey_data_cleaned) +
  geom_boxplot(aes(y=age,x= income, fill=income),
                  position="identity")  +
  scale_fill_brewer(palette="Reds")
ggplot(survey_data_cleaned) +
  geom_boxplot(aes(y=age, x= education, fill=education),
                  position="identity") +
  scale_fill_brewer(palette="Reds")

```

```{r income plots}

# plot income, reorder income brackets
survey_data_cleaned$income <- factor(survey_data_cleaned$income,
                                     levels = c("Less than $15,000",
                                                "$15,000 - $30,000",
                                                "$30,000 - $50,000",
                                                "$50,000 - $75,000",
                                                "$75,000 - $150,000",
                                                "Above $150,000"))
((summary(survey_data_cleaned$income)/length(survey_data_cleaned$income))*100)
# plot looking at income total
ggplot(survey_data_cleaned, aes(x=income, fill=income))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Participant Household Income 2013", y = "Percent", x ='income', fill="income bracket") +
  scale_fill_brewer(palette="Greens")+
  theme_minimal()

# income separated by gender
ggplot(survey_data_cleaned, aes(x=income, group=gender))+
  geom_bar(aes(y = ..prop.., fill =factor(..x..)), stat="count") +
  geom_text(aes(label = scales::percent(..prop..),
                y= ..prop..), stat= "count", vjust = -.5) +
  labs(y = "Percent", x ='income', fill="income bracket") +
  facet_grid(~gender) +
  scale_y_continuous(labels = scales::percent)+
  scale_fill_brewer(palette="Reds")+
  theme_minimal()


# income separated by age
ggplot(survey_data_cleaned, aes(x=income, y=age, fill=income)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Reds")

```

```{r education plots}

# plot education, reorder education values
survey_data_cleaned$education <- factor(survey_data_cleaned$education,
                                     levels = c("Other",
                                                "Less than 12 years; no high school diploma",
                                                "High school graduate (or equivalent)",
                                                "Some college, no diploma; or Associate’s degree",
                                                "Bachelor’s degree or other 4-year degree",
                                                "Graduate or professional degree"))
((summary(survey_data_cleaned$education)/length(survey_data_cleaned$education))*100)
# plot looking at education
ggplot(survey_data_cleaned, aes(x=education, fill=education))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Survey Highest Degree", y = "Percent", x ='Highest Degree') +
  scale_fill_brewer(palette="Greens")+
  theme_minimal() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# education separated by gender
ggplot(survey_data_cleaned, aes(x=education, group=gender))+
  geom_bar(aes(y = ..prop.., fill =factor(..x..)), stat="count") +
  geom_text(aes(label = scales::percent(..prop..),
                y= ..prop..), stat= "count", vjust = -.5) +
  labs(title = "Survey Highest Degree", y = "Percent", x ='Highest Degree') +
  facet_grid(~gender) +
  scale_y_continuous(labels = scales::percent)+
  scale_fill_brewer(palette="Reds")+
  theme_minimal()

# education separated by age
ggplot(survey_data_cleaned, aes(x=education, y=age, fill=education)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Reds")

```


```{r age worktime}

summary(survey_data_cleaned$worktime)

# plot looking at worktime total
ggplot(survey_data_cleaned) +
  geom_histogram( aes(x=worktime, y = ..density..),
                  bins=30, fill= "lightgreen", position="identity") 

ggplot(survey_data_cleaned) +
  geom_boxplot(aes(y=worktime, x= gender, fill=gender),
                  position="identity")

ggplot(survey_data_cleaned) +
  geom_boxplot(aes(y=worktime,x= income, fill=income),
                  position="identity") 
ggplot(survey_data_cleaned) +
  geom_boxplot(aes(y=worktime, x= education, fill=education),
                  position="identity") 

```

```{r sirius plots}


# plot looking at sirius listeners
ggplot(survey_data_cleaned, aes(x=sirius, fill=sirius))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Self repoted listeners to SiriusXM", y = "Percent", x ='Answer') +
  theme_minimal() +
  scale_fill_brewer(palette="Greens")


survey_data_cleaned[survey_data_cleaned$sirius =='Yes',] %>% 
  ggplot(., aes(x=gender, fill = gender)) +
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Self repoted listeners to SiriusXM", y = "Percent", x ='Answer') +
  theme_minimal() +
  scale_fill_brewer(palette="Reds")

survey_data_cleaned[survey_data_cleaned$sirius =='Yes',] %>% 
  ggplot(., aes(x=education, fill=education))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Self repoted listeners to SiriusXM", y = "Percent", x ='Answer') +
  theme_minimal() +
  scale_fill_brewer(palette="Reds")

survey_data_cleaned[survey_data_cleaned$sirius =='Yes',] %>% 
  ggplot(., aes(x=income, fill=income))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Self repoted listeners to SiriusXM", y = "Percent", x ='Answer') +
  theme_minimal() +
  scale_fill_brewer(palette="Reds")

```

```{r wharton plots}

ggplot(survey_data_cleaned, aes(x=wharton, fill=wharton))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Self repoted listeners to wharton", y = "Percent", x ='Answer') +
  theme_minimal() +
  scale_fill_brewer(palette="Greens")

survey_data_cleaned[survey_data_cleaned$wharton =='Yes',] %>% 
  ggplot(., aes(x=gender, fill = gender)) +
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Self repoted listeners to SiriusXM", y = "Percent", x ='Answer') +
  theme_minimal() +
  scale_fill_brewer(palette="Reds")

survey_data_cleaned[survey_data_cleaned$wharton =='Yes',] %>% 
  ggplot(., aes(x=education, fill=education))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Self repoted listeners to SiriusXM", y = "Percent", x ='Answer') +
  theme_minimal() +
  scale_fill_brewer(palette="Reds")

survey_data_cleaned[survey_data_cleaned$wharton =='Yes',] %>% 
  ggplot(., aes(x=income, fill=income))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Self repoted listeners to SiriusXM", y = "Percent", x ='Answer') +
  theme_minimal() +
  scale_fill_brewer(palette="Reds")
```

## Sample properties

The population from which the sample is drawn determines where the results of our analysis can be applied or generalized. We include some basic demographic information for the purpose of identifying sample bias, if any exists. Combine our data and the general population distribution in age, gender and income to try to characterize our sample on hand.

1. Does this sample appear to be a random sample from the general population of the USA?

2. Does this sample appear to be a random sample from the MTURK population?

Note: You can not provide evidence by simply looking at our data here. For example, you need to find distribution of education in our age group in US to see if the two groups match in distribution. You may need to gather some background information about the MTURK population to have a slight sense if this particular sample seem to a random sample from there... Please do not spend too much time gathering evidence. 

from the Census Bureau's Annual Estimates of the Resident Population by Sex, Race, and Hispanic Origin for the United States: April 1, 2010 to July 1, 2019	

```{r US census Bureau data: sex}
# US census bureau data
female_count <- 161646584
male_count <- 156654424
us_gender_2014 <- data.frame("gender" = c("Male", "Female"),
                             "pct" = c(male_count/(male_count+female_count),
                                       female_count/(male_count+female_count)))


# plot looking at gender
ggplot(survey_data_cleaned, aes(x=gender, fill=gender))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Participants by Gender", y = "Percent", x ='Gender') +
  theme_minimal() +
  scale_fill_brewer(palette="Greens")

ggplot(us_gender_2014, aes(x=gender, y=pct, fill=gender))+
  geom_bar(stat= "identity")  +
  geom_text(aes(label = scales::percent(pct),
                y= pct, stat= "count", vjust = -.5)) +
  labs(title = "US census by Gender 2014", y = "Percent", x ='Gender') +
  theme_minimal() +
  scale_fill_brewer(palette="Greens")

```

```{r US census Bureau data: age}
# US census bureau data

us_census_age_values <- read_excel("data/nc-est2019-agesex.xlsx", range = "T7:T24", col_names = FALSE)
us_census_age_categories <- gsub('.','',
                                 read_excel("data/nc-est2019-agesex.xlsx",
                                            range = "A7:A24", col_names = FALSE)[[1]],
                                 fixed = TRUE)
us_census_age <- data.frame("ranges" = gsub('.','',us_census_age_categories, fixed = TRUE),
                       "count" = us_census_age_values[[1]])
head(us_census_age)
names(us_census_age)
us_census_age$ranges

survey_age <- data.frame("age" = c("Under 5 years" = 0,
                                         "5 to 9 years"= 0,
                                         "10 to 14 years"= 0,
                                         "15 to 19 years" = sum(survey_data_cleaned$age < 19),
                                         "20 to 24 years" = sum(survey_data_cleaned$age > 19 &
                                                                survey_data_cleaned$age < 25),
                                         "25 to 29 years" = sum(survey_data_cleaned$age > 24 &
                                                                survey_data_cleaned$age < 30),
                                         "30 to 34 years" = sum(survey_data_cleaned$age > 29 &
                                                                survey_data_cleaned$age < 35),
                                         "35 to 39 years" = sum(survey_data_cleaned$age > 34 &
                                                                survey_data_cleaned$age < 40),
                                         "40 to 44 years" = sum(survey_data_cleaned$age > 39 &
                                                                survey_data_cleaned$age < 45),
                                         "45 to 49 years" = sum(survey_data_cleaned$age > 44 &
                                                                survey_data_cleaned$age < 50),
                                         "50 to 54 years"= sum(survey_data_cleaned$age > 49 &
                                                                survey_data_cleaned$age < 55),
                                         "55 to 59 years" = sum(survey_data_cleaned$age > 54 &
                                                                survey_data_cleaned$age < 60),
                                         "60 to 64 years" = sum(survey_data_cleaned$age > 59 &
                                                                survey_data_cleaned$age < 65),
                                         "65 to 69 years" = sum(survey_data_cleaned$age > 64 &
                                                                survey_data_cleaned$age < 70),
                                         "70 to 74 years" = sum(survey_data_cleaned$age > 69 &
                                                                survey_data_cleaned$age < 75),
                                         "75 to 79 years" = sum(survey_data_cleaned$age > 74 &
                                                                survey_data_cleaned$age < 80),
                                         "80 to 84 years" = sum(survey_data_cleaned$age > 79 &
                                                                survey_data_cleaned$age < 85),
                                         "85 years and over" = sum(survey_data_cleaned$age > 84)))

survey_age <- data.frame("ranges" = rownames(survey_age), "count"=survey_age$age)

survey_age$ranges <- factor(survey_age$ranges,
                                     levels = c("Under 5 years","5 to 9 years",
                                                "10 to 14 years","15 to 19 years",
                                                "20 to 24 years","25 to 29 years",
                                                "30 to 34 years","35 to 39 years",
                                                "40 to 44 years","45 to 49 years",
                                                "50 to 54 years","55 to 59 years",
                                                "60 to 64 years","65 to 69 years",
                                                "70 to 74 years","75 to 79 years",
                                                "80 to 84 years","85 years and over"))
us_census_age$ranges <- factor(us_census_age$ranges,
                                     levels = c("Under 5 years","5 to 9 years",
                                                "10 to 14 years","15 to 19 years",
                                                "20 to 24 years","25 to 29 years",
                                                "30 to 34 years","35 to 39 years",
                                                "40 to 44 years","45 to 49 years",
                                                "50 to 54 years","55 to 59 years",
                                                "60 to 64 years","65 to 69 years",
                                                "70 to 74 years","75 to 79 years",
                                                "80 to 84 years","85 years and over"))

ggplot(survey_age, aes(x=ranges, y=count, fill = ranges))+
  geom_bar(stat= "identity")  +
  labs(title = "Survey Age", y = "Percent", x ='Age Groups') +
  theme_minimal()

ggplot(us_census_age, aes(x=ranges, y=count, fill = ranges))+
  geom_bar(stat= "identity")  +
  labs(title = "US Census Age", y = "Percent", x ='Age Groups') +
  theme_minimal()
```

```{r US census Bureau data: education}
# US census bureau data

us_census_education <- read_excel("data/table-1-01.xlsx", range = "C6:Q21")[-1,]
names(us_census_education)
us_education_2014 <- data.frame("education" = c(rep("Other", 0),
                                             rep("Less than 12 years; no high school diploma",
                                                 sum(us_census_education[,c(1:7)])),
                                             rep("High school graduate (or equivalent)",
                                                 sum(us_census_education$`High school graduate`)),
                                             rep("Some college, no diploma; or Associate’s degree",
                                                 sum(us_census_education[,c(9:11)])),
                                             rep("Bachelor’s degree or other 4-year degree",
                                                 sum(us_census_education$`Bachelor's degree`)),
                                             rep("Graduate or professional degree",
                                                 sum(us_census_education[,c(13:15)]))))

us_education_2014$education <- factor(us_education_2014$education,
                                     levels = c("Other",
                                                "Less than 12 years; no high school diploma",
                                                "High school graduate (or equivalent)",
                                                "Some college, no diploma; or Associate’s degree",
                                                "Bachelor’s degree or other 4-year degree",
                                                "Graduate or professional degree"))

ggplot(survey_data_cleaned, aes(x=education, fill=education))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Survey Highest Degree", y = "Percent", x ='Highest Degree') +
  scale_fill_brewer(palette="Greens")+
  theme_minimal() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

ggplot(us_education_2014, aes(x=education, fill=education))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "US Education 2014", y = "Percent", x ='Highest Degree', fill="income bracket") +
  scale_fill_brewer(palette="Greens")+
  theme_minimal()
```

```{r US census Bureau data: Income}
# US census bureau data
# https://www.census.gov/data/tables/time-series/demo/income-poverty/cps-hinc/hinc-01.2013.html#list-tab-HKAIF65SA26ZC3YMQB
us_income_2013_excel <- read_excel("data/hinc01_1.xls", range = "C8:AQ10")[2,]
names(us_income_2013_excel)
us_income_2013 <- data.frame("income" = c(rep("Less than $15,000", sum(us_income_2013_excel[,c(1:3)])),
                                          rep("$15,000 - $30,000", sum(us_income_2013_excel[,c(4:6)])),
                                          rep("$30,000 - $50,000", sum(us_income_2013_excel[,c(7:10)])),
                                          rep("$50,000 - $75,000", sum(us_income_2013_excel[,c(11:15)])),
                                          rep("$75,000 - $150,000", sum(us_income_2013_excel[,c(16:30)])),
                                          rep("Above $150,000", sum(us_income_2013_excel[,c(30:41)]))))

us_income_2013$income <- factor(us_income_2013$income,
                                     levels = c("Less than $15,000",
                                                "$15,000 - $30,000",
                                                "$30,000 - $50,000",
                                                "$50,000 - $75,000",
                                                "$75,000 - $150,000",
                                                "Above $150,000"))

ggplot(survey_data_cleaned, aes(x=income, fill=income))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "Survey Household Income 2013", y = "Percent", x ='income', fill="income bracket") +
  scale_fill_brewer(palette="Greens")+
  theme_minimal()

ggplot(us_income_2013, aes(x=income, fill=income))+
  geom_bar(aes(y = (..count..)/sum(..count..)))  +
  geom_text(aes(label = scales::percent((..count..)/sum(..count..)),
                y = ((..count..)/sum(..count..))), stat= "count", vjust = -.5) +
  labs(title = "US Household Income 2013", y = "Percent", x ='income', fill="income bracket") +
  scale_fill_brewer(palette="Greens")+
  theme_minimal()
```


## Final estimate

Give a final estimate of the Wharton audience size in January 2014. Assume that the sample is a random sample of the MTURK population, and that the proportion of Wharton listeners vs. Sirius listeners in the general population is the same as that in the MTURK population. Write a brief executive summary to summarize your findings and how you came to that conclusion.

```{r estimating population size}
total_sirius_liteners <- survey_data_cleaned[survey_data_cleaned$sirius == "Yes",]
total_wharton_liteners <- total_sirius_liteners[total_sirius_liteners$wharton == "Yes",]

prop_listeners <- nrow(total_wharton_liteners)/nrow(total_sirius_liteners)
prop_listeners

51600000 * prop_listeners
```

To be specific, you should include:

1. Goal of the study
2. Method used: data gathering, estimation methods
3. Findings
4. Limitations of the study. 


## New task

Now suppose you are asked to design a study to estimate the audience size of Wharton Business Radio Show as of today: You are given a budget of $1000. You need to present your findings in two months. 

Write a proposal for this study which includes:

1. Method proposed to estimate the audience size.
2. What data should be collected and where it should be sourced from.
Please fill in the google form to list your platform where surveys will be launched and collected [HERE](https://forms.gle/8SmjFQ1tpqr6c4sa8) 

A good proposal will give an accurate estimation with the least amount of money used. 





# Case study 2: Women in Science


Are women underrepresented in science in general? How does gender relate to the type of educational degree pursued? Does the number of higher degrees increase over the years? In an attempt to answer these questions, we assembled a data set (`WomenData_06_16.xlsx`) from [NSF](https://ncses.nsf.gov/pubs/nsf19304/digest/field-of-degree-women) about various degrees granted in the U.S. from 2006 to 2016. It contains the following variables: Field (Non-science-engineering (`Non-S&E`) and sciences (`Computer sciences`, `Mathematics and statistics`, etc.)), Degree (`BS`, `MS`, `PhD`), Sex (`M`, `F`), Number of degrees granted, and Year.

Our goal is to answer the above questions only through EDA (Exploratory Data Analyses) without formal testing. We have provided sample R-codes in the appendix to help you if needed. 


## Data preparation  

1. Understand and clean the data


Notice the data came in as an Excel file. We need to use the package `readxl` and the function `read_excel()` to read the data `WomenData_06_16.xlsx` into R. 


a). Read the data into R.

b). Clean the names of each variables. (Change variable names to  `Field`,`Degree`, `Sex`, `Year` and `Number` )

c). Set the variable natures properly. 

d). Any missing values?

```{r data wrangling, echo = FALSE, warning = FALSE}
wsci <- read_excel("data/WomenData_06_16.xlsx")
names(wsci)
head(wsci)
summary(wsci)
# change names
# set the field, degree and sex as factors
wsci %<>% rename(Field = 'Field and sex',
                 Number = "Degrees Awarded") %>% 
  mutate(Field = as.factor(Field),
         Sex = as.factor(Sex),
         Degree = as.factor(Degree))
```

2. Write a summary describing the data set provided here. 

a). How many fields are there in this data?
5 fields: Field, Degree, Sex, Year, Number

```{r summary, echo = FALSE, warning = FALSE}
str(wsci)

summary(wsci$Field)
summary(wsci$Degree)
summary(wsci$Sex)
summary(wsci$Year)
unique(wsci$Year)
length(unique(wsci$Year))
summary(wsci$Number)
```

b). What are the degree types? 
BS MS PhD 

c). How many year's statistics are being reported here? 
11


## BS degrees in 2015

Is there evidence that more males are in science-related fields vs `Non-S&E`? Provide summary statistics and a plot which shows the number of people by gender and by field. Write a brief summary to describe your findings.


```{r}

# by field (S&E and Non-S&E) only
wsci %>%
  mutate(SE = ifelse(Field!="Non-S&E" , "S&E", "Non-S&E")) %>%
  group_by(SE) %>%
  summarise(SE_number = sum(Number)) %>%
  ggplot(aes(x = SE, y = SE_number, fill = SE)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =SE_number), vjust = -.5)  +
  ggtitle("Degrees Awarded by Field (S&E and Non-S&E)") 

# by sex and field (S&E and Non-S&E)
wsci %>%
  mutate(SE = ifelse(Field!="Non-S&E" , "S&E", "Non-S&E")) %>%
  group_by(SE, Sex) %>%
  summarise(SE_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = SE_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =SE_number), vjust = -.5)  +
  facet_wrap(~SE, scales = "free_y") +
  ggtitle("Degrees Awarded by Sex and Field (S&E and Non-S&E)") 


```

## EDA bringing type of degree, field and gender in 2015

Describe the number of people by type of degree, field, and gender. Do you see any evidence of gender effects over different types of degrees? Again, provide graphs to summarize your findings.

```{r eval = FALSE, echo = FALSE}
# number of degrees by sex
wsci %>%
  group_by(Sex) %>%
  summarise(sex_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = sex_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =sex_number), vjust = -.5)  +
  ggtitle("Degrees Awarded by Sex") 

# number of degrees by degree
wsci %>%
  group_by(Degree) %>%
  summarise(degree_number = sum(Number)) %>%
  ggplot(aes(x = Degree, y = degree_number, fill = Degree)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =degree_number), vjust = -.5)  +
  ggtitle("Degrees Awarded by Degree Type") 

# number of degrees by field
wsci %>%
  group_by(Field) %>%
  summarise(field_number = sum(Number)) %>%
  ggplot(aes(x = Field, y = field_number, fill = Field)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =field_number), vjust = -.5)  +
  ggtitle("Degrees Awarded by Field") 

# number of degrees by degree and sex
wsci %>%
  group_by(Degree, Sex) %>%
  summarise(degree_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = degree_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =degree_number), vjust = -.5)  +
  facet_wrap(~Degree, scales = "free_y") +
  ggtitle("Degrees Awarded Sex and Degree Type") 

# number of degrees by field and sex
wsci %>%
  group_by(Field, Sex) %>%
  summarise(field_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = field_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =field_number), vjust = -.5)  +
  facet_wrap(~Field, scales = "free_y") +
  ggtitle("Degrees Awarded By Sex and Field") 

# number of degrees by field and degree
wsci %>%
  group_by(Field, Degree) %>%
  summarise(field_number = sum(Number)) %>%
  ggplot(aes(x = Field, y = field_number, fill =Field)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =field_number), vjust = -.5)  +
  facet_wrap(~Degree, scales = "free_y") +
  ggtitle("Degrees Awarded by Field and Degree Type") 


# number of degrees by sex and field and degree
wsci %>%
  group_by(Field, Sex, Degree) %>%
  summarise(All_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = All_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(Degree~Field, scales = "free_y") +
  geom_text(aes(label =All_number), vjust = -.5)  +
  ggtitle("Degrees Awarded by Sex, Field, and Degree Type") 

```

## EDA bring all variables 

In this last portion of the EDA, we ask you to provide evidence numerically and graphically: Do the number of  degrees change by gender, field, and time? 

```{r eval = FALSE, echo = FALSE}

# time series: degrees awarded by field
wsci %>%
  group_by(Field, Year) %>%
  summarise(field_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = field_number, color = Field)) +
  geom_point() + geom_line() +
  ggtitle("Degrees awarded across years by Field")

# time series: degrees awarded by sex
wsci %>%
  group_by(Sex, Year) %>%
  summarise(sex_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = sex_number, color = Sex)) +
  geom_point() + geom_line() +
  ggtitle("Degrees awarded across years by Sex")

# time series: degrees awarded by degree type
wsci %>%
  group_by(Degree, Year) %>%
  summarise(degree_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = degree_number, color = Degree)) +
  geom_point() + geom_line() +
  ggtitle("Degrees awarded across years by Degree")

# time series: degrees awarded by Field and sex
wsci %>%
  group_by(Field, Sex, Year) %>%
  summarise(field_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = field_number, color = Sex)) +
  geom_point() + geom_line() +
  facet_wrap(~Field)+
  ggtitle("Degrees awarded across years by Degree")

# time series: degrees awarded by Field and Degree
wsci %>%
  group_by(Field, Degree, Year) %>%
  summarise(field_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = field_number, color = Degree)) +
  geom_point() + geom_line() +
  facet_wrap(~Field)+
  ggtitle("Degrees awarded across years by Degree")

# time series: degrees awarded by Sex and Degree
wsci %>%
  group_by(Sex, Degree, Year) %>%
  summarise(sex_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = sex_number, color = Sex)) +
  geom_point() + geom_line() +
  facet_wrap(~Degree)+
  ggtitle("Degrees awarded across years by Degree")

# time series: degrees awarded by Sex and Field and Degree
wsci %>%
  group_by(Sex, Degree, Field, Year) %>%
  summarise(sex_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = sex_number, color = Sex)) +
  geom_point() + geom_line() +
  facet_grid(Field~Degree)+
  ggtitle("Degrees awarded across years by Degree")
```

## Women in Data Science

Finally, is there evidence showing that women are underrepresented in data science? Data science is an interdisciplinary field of computer science, math, and statistics. You may include year and/or degree.

```{r eval = FALSE, echo = FALSE}
## by sex, field (Social Sciences and Psychology), degree, and time

# sex and field
wsci %>%
  filter(Field %in% c("Computer sciences", "Mathematics and statistics")) %>%
  group_by(Field, Sex) %>%
  summarise(All_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = All_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =All_number), vjust = -.5)  +
  facet_grid(~Field, scales = "free_y") +
  ggtitle("Degrees granted in Data science by sex")

# Sex, Field, Degree
wsci %>%
  filter(Field %in% c("Computer sciences", "Mathematics and statistics")) %>%
  group_by(Field, Sex, Degree) %>%
  summarise(All_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = All_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label =All_number), vjust = -.5)  +
  facet_grid(Degree~Field, scales = "free_y") +
  ggtitle("Degrees granted in Data science by sex")

# Sex, Field, Degree, Time
wsci %>%
  filter(Field %in% c("Computer sciences", "Mathematics and statistics")) %>%
  group_by(Field, Sex, Degree, Year) %>%
  summarise(All_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = All_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(Degree~Field, scales = "free_y") +
  ggtitle("Degrees granted in Data science by sex")

```

```{r}
# social science

# Sex, Field
wsci %>%
  filter(Field %in% c("Psychology", "Social sciences")) %>%
  group_by(Field, Sex) %>%
  summarise(All_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = All_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(~Field, scales = "free_y") +
  ggtitle("Degrees granted in Social science by sex across degree and SE")

# Sex, Field, Degree
wsci %>%
  filter(Field %in% c("Psychology", "Social sciences")) %>%
  group_by(Field, Sex, Degree) %>%
  summarise(All_number = sum(Number)) %>%
  ggplot(aes(x = Sex, y = All_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(Field~Degree, scales = "free_y") +
  ggtitle("Degrees granted in Social science by sex across degree and SE")

# Sex, Field, Degree, Time
wsci %>%
  filter(Field %in% c("Psychology", "Social sciences")) %>%
  group_by(Field, Sex, Degree,Year) %>%
  summarise(All_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = All_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(Field~Degree, scales = "free_y") +
  ggtitle("Degrees granted in Social science by sex across degree and SE")


```

## Final brief report

Summarize your findings focusing on answering the questions regarding if we see consistent patterns that more males pursue science-related fields. Any concerns with the data set? How could we improve on the study?

## Appendix

To help out, we have included some R-codes here as references. You should make your own chunks filled with texts going through each items listed above. Make sure to hide the unnecessary outputs/code etc. 

1. Clean data

```{r data wrangling, echo = FALSE, warning = FALSE}
# For the demonstration purpose, we show this R-chunk by taking echo=TRUE
# In your final report you should hide all the R-chunks to keep your report flowing well.
wsci <- read_excel("WomenData_06_16.xlsx")
names(wsci)
head(wsci)
#change names
wsci %<>% 
  rename(Field = 'Field and sex')
# set the field, degree and sex as factors
wsci %<>% 
  mutate( Field = as.factor(Field))
```


```{r, echo=FALSE}
# wsci %<>%
#   rename(Field = "Field and sex",
#          Number = "Degrees Awarded") %>%
#   mutate(Field = as.factor(Field),
#          Degree = as.factor(Degree),
#          Sex = as.factor(Sex))

```


2. A number of sample analyses 

```{r eval = FALSE, echo = FALSE}
wsci %>%  # to get the average number of ppl by gender
  group_by(Field, Sex) %>%
  summarise(deg = mean(Number))

wsci %>%
  filter(Year == 2007) %>%
  ggplot(aes(x = Field, y = Number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(Degree~., scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  ggtitle("Degrees granted across fields by degree and gender") 

wsci %>%
  mutate(SE = ifelse(Field!="Non-S&E" , "S&E", "Non-S&E")) %>%
  group_by(SE, Sex) %>%
  summarise(SE_number = sum(Number)) %>%
  ggplot(aes(x = SE, y = SE_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.y = element_text(angle = 60)) +
  ggtitle("Degrees granted by S&E vs non-S&E by gender")

wsci %>%
  mutate(SE = ifelse(Field!="Non-S&E" , "S&E", "Non-S&E")) %>%
  group_by(SE, Sex, Year) %>%
  summarise(SE_number = sum(Number)) %>%
  group_by(SE, Year) %>%
  mutate(ratio = SE_number / sum(SE_number)) %>%
  filter(Sex == "Female") %>%
  ggplot(aes(x = Year, y = ratio, color = SE)) +
  geom_point() + geom_line() +
  ggtitle("Female proportion in SE/non-SE across year")

wsci %>%
  mutate(SE = ifelse(Field!="Non-S&E" , "S&E", "Non-S&E")) %>%
  group_by(SE, Sex, Year, Degree) %>%
  summarise(SE_number = sum(Number)) %>%
  group_by(SE, Year, Degree) %>%
  mutate(ratio = SE_number / sum(SE_number)) %>%
  filter(Sex == "Female") %>%
  ggplot(aes(x = Year, y = ratio, color = SE)) +
  geom_point() + geom_line() +
  facet_grid(~Degree)+
  ggtitle("Female proportion in SE/non-SE across year by degree")


wsci %>%
  mutate(SE = ifelse(Field!="Non-S&E" , "S&E", "Non-S&E")) %>%
  group_by(SE, Sex, Year, Degree) %>%
  summarise(SE_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = SE_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(SE~Degree, scales = "free_y") +
  ggtitle("Degrees granted by sex, degree and SE")


wsci %>%
  mutate(SE = ifelse(Field!="Non-S&E" , "S&E", "Non-S&E")) %>%
  group_by(SE, Sex, Year, Degree) %>%
  summarise(SE_number = sum(Number)) %>%
  ggplot(aes(x = Year, y = SE_number, fill = Sex)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(SE~Degree, scales = "free_y") +
  ggtitle("Degrees granted proportion by sex across degree and SE")


wsci %>%
  filter(Field %in% c("Computer sciences", "Mathematics and statistics")) %>%
  ggplot(aes(x = Year, y = Number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(Field~Degree, scales = "free_y") +
  ggtitle("Degrees granted pr option by sex across degree and SE")


wsci %>%
  ggplot(aes(x = Year, y = Number, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(Field~Degree, scales = "free_y") +
  ggtitle("Degrees granted proportion by sex across degree and SE")
```




# Case study 3: Major League Baseball

We would like to explore how payroll affects performance among Major League Baseball teams. The data is prepared in two formats record payroll, winning numbers/percentage by team from 1998 to 2014. 

Here are the datasets:

-`MLPayData_Total.csv`: wide format
-`baseball.csv`: long format

Feel free to use either dataset to address the problems. 

```{r read data}
datapay <- read.csv("data/MLPayData_Total.csv", header=T, stringsAsFactors = FALSE) 
head(datapay)

baseball <- read.csv("data/baseball.csv", header = TRUE, stringsAsFactors = F)
names(baseball)
str(baseball)
summary(baseball)
baseball %>% arrange(forcats::fct_reorder(team, -win_pct, .fun = max))
```

## EDA: Relationship between payroll changes and performance

Payroll may relate to performance among ML Baseball teams. One possible argument is that what affects this year's performance is not this year's payroll, but the amount that payroll increased from last year. Let us look into this through EDA. 

Create increment in payroll

a). To describe the increment of payroll in each year there are several possible approaches. Take 2013 as an example:

    - option 1: diff: payroll_2013 - payroll_2012
    - option 2: log diff: log(payroll_2013) - log(payroll_2012)

Explain why the log difference is more appropriate in this setup.

b). Create a new variable `diff_log=log(payroll_2013) - log(payroll_2012)`. Hint: use `dplyr::lag()` function.

```{r}
baseball %>%
  group_by(team) %>%
  mutate(diff_log =log(payroll) - log(lag(payroll, default = payroll[1], order_by = team))) %>%
  head()
```

c). Create a long data table including: team, year, diff_log, win_pct
```{r}
baseball_long <- baseball %>%
  group_by(team) %>%
  mutate(diff_log =log(payroll) - log(lag(payroll, default = payroll[1], order_by = team))) %>%
  select(., c("team", "year", "diff_log", "win_pct"))
head(baseball_long)
```

## Exploratory questions

a). Which five teams had highest increase in their payroll between years 2010 and 2014, inclusive?

```{r}
baseball_long %>% 
  filter(year >= 2010 & year <=2014) %>%
  arrange(desc(diff_log)) %>% head()
```

b). Between 2010 and 2014, inclusive, which team(s) "improved" the most? That is, had the biggest percentage gain in wins?
```{r}
baseball_long %>%
  group_by(team) %>%
  mutate(diff_log_win =log(win_pct) - log(lag(win_pct, default = win_pct[1], order_by = team))) %>% 
  filter(year >= 2010 & year <=2014) %>%
  arrange(desc(diff_log_win)) %>% head()

```

## Do log increases in payroll imply better performance? 

```{r}
baseball_long <- baseball %>%
  group_by(team) %>%
  mutate(diff_log =log(payroll) - log(lag(payroll, default = payroll[1], order_by = team))) %>%
  select(., c("team", "year", "diff_log", "win_pct")) %>%
  ggplot(aes(x = diff_log, y = win_pct)) + 
  geom_point(aes(color = team), size = 3) + 
  stat_poly_line() +
  stat_poly_eq(aes(label = paste(after_stat(eq.label),
                                 after_stat(rr.label), sep = "*\", \"*"))) +
  labs(title = "MLB Team's Win Percentage  vs. Payroll increase", 
       x = "Log yearly Payroll Increase", 
       y = "Win Percentage") +
  theme_bw() +
  theme(legend.position = "none")


baseball_long <- baseball %>%
  group_by(team) %>%
  mutate(diff_log =log(payroll) - log(lag(payroll, default = payroll[1], order_by = team))) %>%
  select(., c("team", "year", "diff_log", "win_pct")) %>%
  ggplot(aes(x = diff_log, y = win_pct)) + 
  geom_point(aes(color = team), size = 3) + 
  stat_poly_line() +
  stat_poly_eq(aes(label = paste(after_stat(eq.label),
                                 after_stat(rr.label), sep = "*\", \"*"))) +
  labs(title = "MLB Team's Win Percentage  vs. Payroll increase", 
       x = "Log yearly Payroll Increase", 
       y = "Win Percentage") +
  theme_bw() +
  theme(legend.position = "none") +
  facet_wrap(~team)
```

Is there evidence to support the hypothesis that higher increases in payroll on the log scale lead to increased performance?

Pick up a few statistics, accompanied with some data visualization, to support your answer. 

## Comparison

Which set of factors are better explaining performance? Yearly payroll or yearly increase in payroll? What criterion is being used? 






